### 2017-10-15 src/2_Finance.tex
Solo falta relacion con los objetivos en capitulo de conclusiones
fixing table captions

\chapter{High Frequency Trading}

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=\textwidth]{img/HFTtradestime}
  \caption{Holding time of an opened position of a high frequency trade.Source
  \cite{aldridge2009}.}
  \label{fig:HFTtimes}
\end{figure}
In 2014 HFT represented nearly 50\% of equity trades in the US and more than
20\% in Europe showing a consistent fall since 2009 which was its best year. US
market revenues have fallen dramatically.  Figure~\ref{fig:HFTmarket} on the
left shows HFT trades percentage of US and Europe equity shares. The right
figure shows how revenues in US stocks have fallen the last years.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=\textwidth]{img/HFTmarket}
  \caption{Left figure shows HFT market share in the US and Europe. Right figure
  shows revenue in the US. Source: TABB Group.}
  \label{fig:HFTmarket}
\end{figure}
One of the reasons of this fall is that exchange markets have adapted, having
now faster, more transparent and efficient market structures than before.
This has been possible due to its investment in technology enhancing
reliability and stability of transactions. Even though this fall, HFT is still a major component of regulated markets and
will probably remain as a topic of interest for researchers in the near future.

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.7\textwidth]{img/capitalmarkets}
  \caption{Old structure of capital markets. Source \cite{aldridge2009}.}
  \label{fig:capitalmarket}
\end{figure}
Figure~\ref{fig:capitalmarket} shows the typical structure of capital markets
existed from the early 1929s through much of the 1990s where the broker-dealers
played the central and most profitable role.  At the core are the exchanges or
inter-dealer networks (foreign exchange trading). Exchanges are the centralised
marketplaces for transacting.  Broker-dealers perform two functions: trading for
their own accounts and transacting for their customers. Broker-dealers use
inter-dealer brokers to quickly find the best price for a particular asset among
the network of other broker-dealers. Occasionally, broker-dealers also deal
directly with other broker-dealers, particularly for less liquid instruments.
Broker-dealers clients are institutional investors, corporate clients,
commercial clients, and high-net-worth individuals \cite{aldridge2009}.

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/capitalmarketsnow}
  \caption{Actual structure of capital markets. Source \cite{aldridge2009}.}
  \label{fig:capitalmarketnow}
\end{figure}

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/forex-trading-hours.png}
  \caption{Forex market trading hours (GMT). Source \cite{dailyprice}.}
  \label{fig:Forextimes}
\end{figure}

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.5\textwidth]{img/orderbook}
  \caption{Order book. Buyers and sellers are ordered according to the bid or
ask price and the market determines the mid-price and transacted volume. Source
\cite{gould2013limit}.}
  \label{fig:orderbook}
\end{figure}
Today, order books are available and they are a popular source of study among
researchers. The mid-price and spread modelling are some of the
problems related with this area.

>>>>

\chapter{High Frequency Trading}
\label{chapter:HFT}

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=\textwidth]{img/HFTtradestime}
  \caption[Holding time of an opened position of a high frequency trade]{Holding time of an opened position of a high frequency trade.Source
  \cite{aldridge2009}.}
  \label{fig:HFTtimes}
\end{figure}
In 2014 HFT represented nearly 50\% of equity trades in the US and more than
20\% in Europe showing a consistent fall since 2009 which was its best year. US
market revenues have fallen dramatically.  Figure~\ref{fig:HFTmarket} on the
left shows HFT trades percentage of US and Europe equity shares. The right
figure shows how revenues in US stocks have fallen the last years.

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=\textwidth]{img/HFTmarket}
  \caption[HFT market share in the US and Europe]{Left figure shows HFT market share in the US and Europe. Right figure
  shows revenue in the US. Source: TABB Group.}
  \label{fig:HFTmarket}
\end{figure}
One of the reasons of this fall is that exchange markets have adapted, having
now faster, more transparent and efficient market structures than before.
This has been possible due to its investment in technology enhancing
reliability and stability of transactions. Even though this fall, HFT is still a major component of regulated markets and
will probably remain as a topic of interest for researchers in the near future.

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.7\textwidth]{img/capitalmarkets}
  \caption[Old structure of capital markets]{Old structure of capital markets. Source \cite{aldridge2009}.}
  \label{fig:capitalmarket}
\end{figure}
Figure~\ref{fig:capitalmarket} shows the typical structure of capital markets
existed from the early 1929s through much of the 1990s where the broker-dealers
played the central and most profitable role.  At the core are the exchanges or
inter-dealer networks (foreign exchange trading). Exchanges are the centralised
marketplaces for transacting.  Broker-dealers perform two functions: trading for
their own accounts and transacting for their customers. Broker-dealers use
inter-dealer brokers to quickly find the best price for a particular asset among
the network of other broker-dealers. Occasionally, broker-dealers also deal
directly with other broker-dealers, particularly for less liquid instruments.
Broker-dealers clients are institutional investors, corporate clients,
commercial clients, and high-net-worth individuals \cite{aldridge2009}.

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/capitalmarketsnow}
  \caption[Actual structure of capital markets]{Actual structure of capital markets. Source \cite{aldridge2009}.}
  \label{fig:capitalmarketnow}
\end{figure}

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.8\textwidth]{img/forex-trading-hours.png}
  \caption[Forex market trading hours (GMT)]{Forex market trading hours (GMT). Source \cite{dailyprice}.}
  \label{fig:Forextimes}
\end{figure}

@@

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
  \includegraphics[width=0.5\textwidth]{img/orderbook}
  \caption[Order book]{Order book. Buyers and sellers are ordered according to the bid or
ask price and the market determines the mid-price and transacted volume. Source
\cite{gould2013limit}.}
  \label{fig:orderbook}
\end{figure}
Today, order books are available and they are a popular source of study among
researchers. The mid-price and spread modelling are some of the
problems related with this area.

### 2017-10-15 src/3_TimeSeries.tex
changing minor errors
Solo falta relacion con los objetivos en capitulo de conclusiones
fixing table captions

\chapter{Financial Time Series}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/spurious1}
  \caption{Two random walks time series $x_{1,t}$ and $x_{2,t}$}
  \label{fig:spurious1}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=\textwidth]{img/coint-example}
  \caption{Cointegration example: two I(1) processes (blue and green) and one I(0) process (red).}
  \label{fig:cointex}
\end{figure}

@@

Transposing each equation of the system (\ref{eq:vecfull}) we can write
the VECM($p$) model in block-matrix form as:
\begin{equation}\label{eq:simpleveceq} %vareq}
\mathbf{B} = 
\mathbf{A} \mathbf{X} + 
\mathbf{E} \, , 
\end{equation}

@@

When $\mathbf{A}$ is singular, solution to equation~(\ref{eq:vareq}) is given
by the ordinary least squares (OLS) method. OLS consists of minimizing the sum
of squared errors or equivalently minimizing the following expression:

>>>>

\chapter{Financial Time Series}
\label{chapter:FTT}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/spurious1}
  \caption[Random walks time series]{Two random walks time series $x_{1,t}$ and $x_{2,t}$}
  \label{fig:spurious1}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=\textwidth]{img/coint-example}
  \caption[Cointegration example]{Cointegration example: two I(1) processes (blue and green) and one I(0) process (red).}
  \label{fig:cointex}
\end{figure}

@@

Transposing each equation of the system (\ref{eq:vecfull}) we can write
the VECM($p$) model in block-matrix form as:
\begin{equation}\label{eq:simpleveceq}
\mathbf{B} = 
\mathbf{A} \mathbf{X} + 
\mathbf{E} \, , 
\end{equation}

@@

When $\mathbf{A}$ is singular, solution to equation~(\ref{eq:simpleveceq}) is given
by the ordinary least squares (OLS) method. OLS consists of minimizing the sum
of squared errors or equivalently minimizing the following expression:

### 2017-10-15 src/4_OnlineLearning.tex
Solo falta relacion con los objetivos en capitulo de conclusiones
fixing table captions

\chapter{Machine Learning Models}
Machine learning is a scientific discipline focused on the development of
algorithms based on learning from examples. This idea has become central to the
design of search engines, robots systems and forecasts applications that process
large data sets. Usually machines designed to forecast financial time series
requires long training period and therefore they are not suitable to be updated
with stream data. Online machine learning techniques tackle this problem by
updating the model with new data or by computing a new model using less data so
they can give a response in a short period of time.

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/VCdimension}
  \caption{VC dimension example from \cite{vapnik2013nature}. The example shows 2-dimensional vectors $z_k=(x_k,y_k)$,  $k=1,2,3,4$. The left figure
  shows that the 3 dichotomies can be shattered by a line in a plane. However, if we
  add an extra vector $z_4$, it is impossible to shatter them with a
  line. Therefore, linear classifiers has VC dimension equals to $d+1$ where $d$
  is the dimension of the instance space $X$.}
  \label{fig:vcdimension}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/3tradeoff}
  \caption{Tradeoff in empirical learning from \cite{dietterich2003}. The
generalization accuracy is not affected by the amount of examples for low
complex classifiers. However, for some more complex classifiers, the number of
examples improves the generalization accuracy. The key is to find the best
generalization accuracy and the lowest complex classifier with the available
training data. }
  \label{fig:tradeoff}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/model_complexity}
  \caption{Training and test error. Image source is
\cite{friedman2001elements}.The blue curve represents the training error which
decreases with more complex models. However, the model starts to overfit the
data reducing the generalisation capacity of the model,this occurs when the
prediction error in the test sample starts to increase (red curve).}
  \label{fig:traintesterror}
\end{figure}

@@

\begin{figure}[h!]
\includegraphics[width=\linewidth]{img/biasvariance}
\caption{Bias-variance tradeoff depending on lambda. Dotted line corresponds to OLS prediction error (invariant with lambda). Black line corresponds to RR prediction error which can be decomposed in $Bias^2$ (in red) + Variance (in blue).}
\label{fig:biasvariance}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/ffn}
  \caption{Feed-forward neural network (FFN). FNN consists of an input layer, an output layer and at least one hidden layer between the input and output layer. }
  \label{fig:ffn}
\end{figure}

>>>>

\chapter{Machine Learning Models}
\label{chapter:MLM}
Machine learning is a scientific discipline focused on the development of
algorithms based on learning from examples. This idea has become central to the
design of search engines, robots systems and forecasts applications that process
large data sets. Usually machines designed to forecast financial time series
requires long training period and therefore they are not suitable to be updated
with stream data. Online machine learning techniques tackle this problem by
updating the model with new data or by computing a new model using less data so
they can give a response in a short period of time.

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/VCdimension}
  \caption[VC dimension example]{VC dimension example from \cite{vapnik2013nature}. The example shows 2-dimensional vectors $z_k=(x_k,y_k)$,  $k=1,2,3,4$. The left figure
  shows that the 3 dichotomies can be shattered by a line in a plane. However, if we
  add an extra vector $z_4$, it is impossible to shatter them with a
  line. Therefore, linear classifiers has VC dimension equals to $d+1$ where $d$
  is the dimension of the instance space $X$.}
  \label{fig:vcdimension}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/3tradeoff}
  \caption[Tradeoff in empirical learning ]{Tradeoff in empirical learning from \cite{dietterich2003}. The
generalization accuracy is not affected by the amount of examples for low
complex classifiers. However, for some more complex classifiers, the number of
examples improves the generalization accuracy. The key is to find the best
generalization accuracy and the lowest complex classifier with the available
training data. }
  \label{fig:tradeoff}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/model_complexity}
  \caption[Training and test error]{Training and test error. Image source is
\cite{friedman2001elements}.The blue curve represents the training error which
decreases with more complex models. However, the model starts to overfit the
data reducing the generalisation capacity of the model,this occurs when the
prediction error in the test sample starts to increase (red curve).}
  \label{fig:traintesterror}
\end{figure}

@@

\begin{figure}[h!]
\includegraphics[width=\linewidth]{img/biasvariance}
\caption[Bias-variance tradeoff depending on lambda]{Bias-variance tradeoff depending on lambda. Dotted line corresponds to OLS prediction error (invariant with lambda). Black line corresponds to RR prediction error which can be decomposed in $Bias^2$ (in red) + Variance (in blue).}
\label{fig:biasvariance}
\end{figure}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/ffn}
  \caption[Feed-forward neural network (FFN)]{Feed-forward neural network (FFN). FNN consists of an input layer, an output layer and at least one hidden layer between the input and output layer. }
  \label{fig:ffn}
\end{figure}

### 2017-10-15 src/5_1_Proposal.tex
fixing table captions

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
   \includegraphics[width=0.9\textwidth]{img/51_Fig1}
  \caption{Distribution of the number of cointegration vectors using $p=1$ lags.
  Four possible values for windows size $L$ are shown (2, 6, 10 and 14) hours (1
  hour = 360 data points).}
  \label{fig:hists}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
   \includegraphics[width=0.9\textwidth]{img/51_Fig2}
  \caption{MSE versus the percentage of cointegration considering 1000
  iterations. Optimum windows size $L$ found was 1060. Below MSE versus $L$ shows a
  rapidly decreasing behaviour, founding minimum at PC = 73\% }
  \label{fig:cointvsmse}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\textwidth]{img/51_Fig3}
  \caption{Computing time of sequential and parallel algorithm is shown in the
  upper figure. Speed-up is shown below.}
  \label{fig:extimes}
\end{figure}

>>>>

\begin{figure}[!h]
  %\vspace{-0.8cm}
  \centering
   \includegraphics[width=0.9\textwidth]{img/51_Fig1}
  \caption[Distribution of the number of cointegration vectors using $p=1$ lags]{Distribution of the number of cointegration vectors using $p=1$ lags.
  Four possible values for windows size $L$ are shown (2, 6, 10 and 14) hours (1
  hour = 360 data points).}
  \label{fig:hists}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
   \includegraphics[width=0.9\textwidth]{img/51_Fig2}
  \caption[MSE versus the percentage of cointegration considering 1000
  iterations]{MSE versus the percentage of cointegration considering 1000
  iterations. Optimum windows size $L$ found was 1060. Below MSE versus $L$ shows a
  rapidly decreasing behaviour, founding minimum at PC = 73\% }
  \label{fig:cointvsmse}
\end{figure}

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\textwidth]{img/51_Fig3}
  \caption[Computing time and Speed-up]{Computing time of sequential and parallel algorithm is shown in the
  upper figure. Speed-up is shown below.}
  \label{fig:extimes}
\end{figure}

### 2017-10-15 src/5_2_Proposal.tex
Conclusions first part
fixing table captions

\chapter{An Online Vector Error Correction Model for Exchange Rates Forecasting}

@@

In this study, an Online VEC model (OVECM) is proposed, which optimises how
model parameters are obtained using a sliding window of the most recent data.
OVECM, unlile AVECM, updates the parameters at each step, instead of obtaining
new ones. This proposal also takes advantage of the long-run relationship
between the time series in order to obtain improved execution times.

@@

\section{Results} \label{sec:52results}

@@

\begin{figure}[!h]
  \centering
   \includegraphics[width = 10.5cm]{img/mapes} 
  \caption{In-sample MAPEs example for 50 minutes. The average of them is
  considered to obtain new cointegration vectors.}
  \label{fig:mapes}
 \end{figure}

@@

\begin{figure}[h]
  \vspace{-0.2cm}
  \centering
   \includegraphics[width = 10cm]{img/accuracy} 
  \caption{OVECM forecasting accuracy example for 50 minutes using $L = 1000$ and $p =3$}
  \label{fig:accuracy}
 \end{figure}

>>>>

\chapter{An Online Vector Error Correction Model for Exchange Rates Forecasting}
\label{chapter:proposal2}

@@

In this study, an Online VEC model (OVECM) is proposed, which optimises how
model parameters are obtained using a sliding window of the most recent data.
OVECM, unlike AVECM, updates the parameters at each step, instead of obtaining
new ones. This proposal also takes advantage of the long-run relationship
between the time series in order to obtain improved execution times.

@@

\section{Experimental results} \label{sec:52results}

@@

\begin{figure}[!h]
  \centering
   \includegraphics[width = 10.5cm]{img/mapes} 
  \caption[In-sample MAPEs example for 50 minutes]{In-sample MAPEs example for 50 minutes. The average of them is
  considered to obtain new cointegration vectors.}
  \label{fig:mapes}
 \end{figure}

@@

\begin{figure}[h]
  \vspace{-0.2cm}
  \centering
   \includegraphics[width = 10cm]{img/accuracy} 
  \caption[OVECM forecasting accuracy example]{OVECM forecasting accuracy example for 50 minutes using $L = 1000$ and $p =3$}
  \label{fig:accuracy}
 \end{figure}

### 2017-10-15 src/firmas.tex
fixing table captions

\thispagestyle{empty}
\cleardoublepage%
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\vspace{1cm}
\begin{center}
\includegraphics[scale=0.6]{fig/logo_usm}\\
\cambiaFont{ppl}{\Large \textsc{Examining Committee}}\\
\bigskip
\end{center}
\vspace{2cm}
\begin{tabular}[c]{lc}
{\textsc Prof. Dr. Luis Salinas} & \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad Técnica Federico Santa María} & {Advisor}\\
Chile & \\
& \\
& \\
{\textsc Prof. Dr. } &  \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad Técnica Federico Santa María} & {Committee Member} \\
Chile & \\
& \\
& \\
{Prof. Dr. } & \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{University} & {External Peer Reviewer}\\
{Chile} & \\
\end{tabular}

>>>>

\thispagestyle{empty}
\cleardoublepage%
\bigskip
\bigskip
\bigskip
\bigskip
\bigskip
\vspace{1cm}
\begin{center}
\includegraphics[scale=0.6]{fig/logo_usm}\\
\cambiaFont{ppl}{\Large \textsc{Examining Committee}}\\
\bigskip
\end{center}
\vspace{2cm}
\begin{tabular}[c]{lc}
{\textsc Prof. Dr. Luis Salinas} & \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad T\'ecnica Federico Santa Mar\'ia} & {Advisor}\\
Chile & \\
& \\
& \\
{\textsc Dr. Claudio Torres} &  \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad T\'ecnica Federico Santa Mar\'ia} & {Co-advisor} \\
Chile & \\

& \\
& \\
{\textsc Dr. Werner Kristjanpoller} &  \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad T\'ecnica Federico Santa Mar\'ia} & {Co-advisor} \\
Chile & \\
& \\
& \\
{Prof. Dr. Christian de Peretti} & \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Ecole Centrale de Lyon} & {External Peer Reviewer}\\
{France} & \\
& \\
& \\
{Prof. Dr. Wilfredo Palma} & \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Pontificia Universidad Cat\'olica de Chile} & {External Peer Reviewer}\\
{Chile} & \\
& \\
& \\
{\textsc Prof. Dr. Hern\'an Astudillo} &  \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
{Universidad T\'ecnica Federico Santa Mar\'ia} & { PhD program Director} \\
Chile & \\
%& \\
%& \\
%{\textsc Prof. Dr. Carlos Castro} &  \begin{picture}(150,5) \line(1,0){150} \end{picture}\\
%{Universidad T\'ecnica Federico Santa Mar\'ia} & {Committee Member} \\
%Chile & \\
\end{tabular}
